---
title: "EDA - Wk1"
author: "MCCurcio"
date: "12/22/2020"
output: html_document
---

# Principles of Analytic Graphics

## Rules for building analytical graphics

See: https://github.com/DataScienceSpecialization

See: "Beautiful Evidence", Edward Tufte

[Coursera1](https://www.coursera.org/learn/exploratory-data-analysis/lecture/o967b/principles-of-analytic-graphics)

1. Show Comparisons; 
    - Evidence is always relative to other competing ideas and hypothesis.
    - Always ask, **"Compared to What?"**
    - Compared to Controls vs Treatment...

2. Show causality, mechanisms, explanations, systematic structure(s)
    - What is your causality framework for thinking about the question/situation.
    - Q. How do you think the world is working?
    - Example given by R.P. is symptom free days vs fine particulates in air.

3. Show multivariate data
    - the real world is multivariate
    - Therefore **show as much data on a single plot as you can**
    - Too much data can dilute or confound the message while distributing the data over a variable may show a different function.
    - **Simpson's paradox**
    
>The [Yule–Simpson effect](http://en.wikipedia.org/wiki/Simpson%27s_paradox) tells us that this "is a paradox in probability and statistics, in which a trend that appears in different groups of data disappears when these groups are combined."
    
4. Integrate the evidence
    - Completely integrate words, numbers, images, diagrams
    - *Combine different modes of info into one graphic!*
    - Data graphics should make use of many modes of data presentation
    - Don't let the tool drive the analysis, make you want...
    
5. Describe & document the evidence with appropriate labels, scales, sources, etc.
    - A graphic should tell a **complete** story that is credible.
    - What was the data source?
    - Document the computer code that made the plot TOO!
    
6. Content is King
    - Analytical presentations ultimately stand or fall depending on the quality, relevance and integrity of their content.
    
    
## Exploratory Graphs-Part 1
    
[Coursera2](https://www.coursera.org/learn/exploratory-data-analysis/lecture/ilRAK/exploratory-graphs-part-1)

- Slides for this and other Data Science courses may be found at:
    - https://github.com/DataScienceSpecialization/courses/
- If you care to use them, they must be downloaded as a zip file and viewed locally. 
- This lesson corresponds to `04_ExploratoryAnalysis/exploratoryGraphs`.

### Graphics for "internal use"

1. Why use a graphic?
    - Understand properties
    - Find patterns in data
    - Suggest models strategies
    - 'Debug' analyses
    - Communicate results

2. Exploratory graphics can be:
    - Made quickly
    - Made in large numbers
    - Are there problems?
    - **Goal is for personal understanding**
    - Axes and legends are generally cleaned later!
    - Color and size are primarily used for info.

3. Common plots and summaries
    - **1D**
    - 5-number summaries
    - Boxplots
    - Histograms
    - Density plots
    - Barplots

## Exploratory Graphs-Part 2

[Coursera3](https://www.coursera.org/learn/exploratory-data-analysis/lecture/ilRAK/exploratory-graphs-part-2)

4. Summaries of Data 
    - **2D**
    - Multiple/overlayed 1-D plots (Lattice/ggplot2)
    - Scatterplots
    - Smooth scatterplots
    - **Greater than 2D**
    - Multiple/overlayed 2-D plots
    - Use color, size, shape to add dimensions
    - Spinning plots
    - Actual 3D plots (usefulness is debatable)

## Base Plotting Systems in R

[Coursera4](https://www.coursera.org/learn/exploratory-data-analysis/lecture/MTRt4/plotting-systems-in-r)

- R has 3 core plotting systems: 

1. Base - intuitive model, `plot()`, adjusting margins is difficult
2. Lattice - `library(lattice)`, `xyplot`, `bwplot`...,  panel plots
3. Ggplot2 - "grammar of graphics"

See: Interesting article by Jeff Leek, [why-i-dont-use-ggplot2](https://simplystatistics.org/2016/02/11/why-i-dont-use-ggplot2/)

## Base Plotting System-Part 1

[Coursera5](https://www.coursera.org/learn/exploratory-data-analysis/lecture/GhFPB/base-plotting-system-part-1)

### Important Base graphics parameters

- `las` - orientation of labels
- `bg` - background color
- `mar` - margins size
- `oma` - outer margin size (default = 0)
- `mfrow` - number of plots per row (rows filled 1st)
- `mcol` - number of plots per col (cols filed 1st)

Example: `par("mar")` > gives default = [5.1, 4.1, 4.1, 2.1]

## Base Plotting System-Part 2

[Coursera6](https://www.coursera.org/learn/exploratory-data-analysis/lecture/m4P1I/base-plotting-system-part-2)

Basic plotting functions:  

    - plot  
    - lines  
    - points  
    - pch - plot characters; ex. dots, x, squares,...  
    - title - "main" & "sub"  
    - mtext - Add text to the margins  
    - axis - Add ticks to axis  
    - legend("topright", legend = data, pch = 20)  
    - abline - lines  
    - xlab, ylab, xlim, ylim  

#### EXAMPLE: Multiple base plots
```{r}
library(datasets)

par(mfrow = c(1,3), mar = c(4,4,2,1), oma = c(0,0,2,0))
with(airquality, {
      plot(Wind, Ozone, main = "Ozone v. Wind")
      plot(Solar.R, Ozone, main = "Ozone v. Solar Rad.")
      plot(Temp, Ozone, main = "Ozone v. Temp")
      mtext("Ozone v. Weather in NYC", outer = TRUE)
}
)
```


### Base Plotting [Demonstration]

[Coursera7](https://www.coursera.org/learn/exploratory-data-analysis/lecture/yUFDH/base-plotting-demonstration)

- 4 margins; 1=Bottom, 2=left, 3=top, 4=right
    - par(mar = c(4,4,1,1))
- Add text to inside of plot
    - text(1,2, "label")
- Add linear model
    - fit <- lm(y ~ x)
    - abline(fit)


```{r}
example(points)
```

```{r}
x <- rnorm(100)
y <- x + rnorm(100)
g <- gl(2, 50, labels = c("male", "female"))

plot(x, y, type = "n")
points(x[g == "male"], y[g == "male"], col = "green", pch = 18)
points(x[g == "female"], y[g == "female"], col = "blue", pch = 17)
```


## Graphics Devices in R-Part 1

[Coursera8](https://www.coursera.org/learn/exploratory-data-analysis/lecture/DQ1md/graphics-devices-in-r-part-1)

A graphics device is an output device.

- Screen device, printer
- PDF, .png, .jpg, .svg, etc, file devices
- In Linux <- x11
- ?devices for info.
- **Example of device functions: plot, xyplot, gplot**

An other way to save plots is by opening a device;
1. Explicitly open a device, ex. pdf(file = file)
2. Call a plotting function, plot will not show on screen
3. Annotate plot if needed
4. Close or turn off device `dev.off()`

```{r eval=FALSE, include=TRUE}
pdf(file = "plot.pdf")
with(iris, plot(x, y))
title(main = "Title")
dev.off()
```

## Graphics Devices in R-Part 2

[Coursera9](https://www.coursera.org/learn/exploratory-data-analysis/lecture/W6KQs/graphics-devices-in-r-part-2)

Vector formats
- pdf
- svg
- win.metafile (old windows only)
- postscript

Bitmap format
- png (lossless)
- jpeg (lossy)
- tiff (lossless)
- bmp

### Opening multiple graphics devices

**Ploting can only occur on 1 graphics device at a time.**

- To find active device, USE: `dev.cur()`
- Every open graphics device is assigned an integer >= 2.
- Change active device; `dev.set(int)`

### Copying Plots

- `dev.copy` cp from one device to another
- `dev.copy2pdf` copy to pdf

Example
```{r}
library(datasets)

with(faithful, plot(eruptions, waiting))
title(main = "Old Faithful Data")
dev.copy(png, file = "geyser_plot.png")
dev.off()
```



Then, follow the menus and select the Exploratory Data Analysis course when given the option. For the first part of this course you should complete the following lessons:

1.    Principles of Analytic Graphs
2.    Exploratory Graphs
3.    Graphics Devices in R
4.    Plotting Systems
5.    Base Plotting System


## Work Collected from Swirl/EDA/Exploratory Graphs

head(), dim(), summary(), quantile(), boxplot(), boxplot(ppm, col = "blue"), 
abline(h=12), hist(ppm, col="green"), rug(ppm), info(), 
hist(ppm, col = "green", breaks = 100), abline(v=12, lwd=2),
abline(v = median(ppm), col = "magenta", lwd = 4), names(pollution), table(),
par(mfrow=c(2,1),mar=c(4,4,2,1)), west <- subset(pollution, region == "west")
hist(subset(pollution,region=="west")$pm25, col = "green"),
plot(pollution$latitude, pollution$pm25, col = pollution$region),

## Work Collected from Swirl/EDA/Graphics Devices in R 

?Devices, with(faithful, plot(eruptions, waiting)), dev.cur(), info(),
pdf(file="myplot.pdf"), dev.off(), 

## Work Collected from Swirl/EDA/Plotting Systems

load ‘ggplot2’ & ‘lattice’ & ‘jpeg’
with(data = cars, plot(speed, dist))
xyplot(Life.Exp ~ Income | region, data = state, layout = c(4,1))

## Work Collected from Swirl/EDA/Base Plotting Systems

range(airquality$Ozone, na.rm = TRUE), hist(airquality$Ozone),
boxplot(Ozone ~ Month, data = airquality,
        xlab = "Month", ylab = "Ozone (ppb)", 
        col.axis = "blue", col.lab = "red"),
with(airquality, plot(Wind, Ozone)) ,

length(par()), names(par()), par()$pin, par("pin"),
par("fg") | par('fg') | par()$fg, par("pch"),
may <- subset(airquality, Month==5),
notmay <- subset(airquality, Month != 5),
points(notmay$Wind, notmay$Ozone,col="red",pch=8),
legend("topright", pch = c(17,8), col = c("blue","red"), legend = c("May", "Other Months"))

par(mfrow = c(1,2))
plot(airquality$Wind, airquality$Ozone, main = "Ozone and Wind")
plot(airquality$Ozone, airquality$Solar.R, main = "Ozone and Solar Radiation")

par(mfrow = c(1, 3), mar = c(4, 4, 2, 1), oma = c(0, 0, 2, 0)),
plot(airquality$Solar.R, airquality$Ozone, main = "Ozone and Solar Radiation")
plot(airquality$Wind, airquality$Ozone, main = "Ozone and Wind")
plot(airquality$Temp, airquality$Ozone, main = "Ozone and Temperature")
mtext("Ozone and Weather in New York City", outer=TRUE)

