---
title: Coursera-R programming-Assignment-3
owner: MCurcio
data: 12/9/2020
---

The function should check the validity of its arguments. If an invalid outcome value is passed to rankall,
the function should throw an error via the stop function with the exact message “invalid outcome”. 

IMPORTANT:
- The `num` variable can take 3 values; “best”, “worst”, or an integer indicating the ranking (smaller numbers are better).

- If the number given by num is larger than the number of hospitals in that state, then the function should
return NA.

```{r}
rankall <- function(outcome = "pneumonia", num = 50) {
      library("tidyverse") # Used for '%>%' & 'arrange'
      outcome = "heart attack"
      num = 50
      # Read .csv
      df = read.csv("outcome-of-care-measures.csv",
                    na.strings = "Not Available",
                    skip = 1,
                    header = FALSE,
                    stringsAsFactors = FALSE)
      
      # Retain only hospital, state, ha, hf, p
      df <- df[,c(2,7,11,17,25)]
      # Collect alphabetized list of states
      all_states <- sort(unique(df[, 2]))
      
      # Check for invalid outcome
      if (!(outcome %in% c("heart attack", "heart failure", "pneumonia"))) { 
            stop("invalid outcome") 
      } else { 
          #print("outcome OK") 
          switch(outcome,
                 "heart attack"  = (df <- df[, c(1, 2, 3)]),
                 "heart failure" = (df <- df[, c(1, 2, 4)]),
                 "pneumonia"     = (df <- df[, c(1, 2, 5)]))
      }
      # Create an empty data frame with column names
      ranked_outcomes <- data.frame("hospital" = character(0), "state" = character(0))
      # Prepare list of hospitals from all states
      for (state in all_states) {
             # subset by state 
             subset_df  <-  df[df[,2] == state,] 
             # arrange by day-score
             subset_df <- subset_df %>% arrange(subset_df[,3], subset_df[,1])
             # Choose best from state
             subset_df <- subset_df[1, ] 
             # Save 'best' 1 to df
             ranked_outcomes <- rbind(ranked_outcomes, subset_df)
      }
      # Order ranked_outcomes by score
      ranked_outcomes <- arrange(ranked_outcomes, ranked_outcomes[,3])
      # Return hospital names and states
      df <- ranked_outcomes[, c(1, 2)]
      
      # Categorize "best" | "worst" | Int(num) | stop("invalid num")
      # if (num %in% c("best", "worst")) { 
      #       switch(num,
      #              "best"  = (return(df)),
      #              "worst" = (return(df %>% arrange(desc(row_number())))))
      # }
      # if (!(is.integer(num))) {
      #      stop("invalid number/term")
      # }
      # if (num > 0 | num < 51) {
      #     print("OK")
      # }
}
```

      print("num OK")
             switch(num,
                    1:50  = (df <- df[, c(1, 2)]),
                    "worst" = (df %>% ranked_outcomes(rev)),
                     = (stop("invalid num term")),
                    num > 50 = (stop("invalid num term")))


```{r}
df <- rankall(outcome = "pneumonia",  num = 50)

df
```


```{r}
best("SC", "heart attack")

df
```


```{r}
df <- rankall(file_name = "outcome-of-care-measures.csv",
           state = "RI", 
           outcome = "heart attack")

head(df)
```

```{r}
df <- best(file_name = "outcome-of-care-measures.csv",
           state = "RI", 
           outcome = "heart battack")

head(df)
```















