---
title: "week 2 mysql"
author: "MCCurcio"
date: "12/21/2020"
output: html_document
---

# Getting & Cleaning Data - Week 2
## Reading from MySQL

https://www.coursera.org/learn/data-cleaning/lecture/njjbw/reading-from-mysql

### For Help:

1. [RMySQL vignette pdf](https://cran.r-project.org/web/packages/RMySQL/RMySQL.pdf)
2. [List of commands](pantz.org/software/mysql/mysqlcommands.html)
    - Do not delete, add or join items from ensembl ONLY select.
    - In general be carful with mysql commands
3. [R-bloggers](https://www.r-bloggers.com/2011/08/mysql-and-r/) - Good?

### Connecting & listing db
```{r}
library(RMySQL)
ucscDb <- dbConnect (MySQL(), # other DB can be used...
	                 user = "genome",
	                 host = "genome-mysql.cse.ucsc.edu")
	                 
result <- dbGetQuery(ucscDb, "show databases;") # dbGetQuery is a mysql command

dbDisconnect(ucscDb) # important to disconnect, shoyuld show TRUE
```

```{r}
result
```

### Connecting to hg19 & listing tables
```{r}
hg19 <- dbConnect(MySQL(),
                  user = "genome",
                  db = "hg19", # human genome 19
                  host = "genome-mysql.cse.ucsc.edu")
                  
allTables <- dbListTables(hg19) # list tables
length(allTables)

```

```{r}
allTables[1:5]
```

```{r}
dbListFields(hg19, "affyU133Plus2")
```

```{r}
dbGetQuery(hg19, "select count(*) from affyU133Plus2")
# Counts all the records in table `affy`
```

# Rading from the table
This gives the records
```{r}
affyData <- dbReadTable(hg19, "affyU133Plus2")
head(affyData)
```

How to select a specific subset of big data
```{r message=FALSE, warning=FALSE}
query <- dbSendQuery(hg19,
                        "select * from affyU133Plus2 where misMatches between 1 and 3")
affyMis <- fetch(query)
quantile(affyMis$misMatches)
```

```{r}
affyMisSmall <- fetch(query, n = 10) # Fecth only top 10 records
dbClearResult(query) # after Send FOR Query one needs to clear the resultant, returning TRUE if it completes.
```

```{r}
dim(affyMisSmall)
```

```{r}
dbDisconnect(hg19) # important to disconnect DB, should show TRUE
```
### See:

1. [RMySQL vignette pdf](https://cran.r-project.org/web/packages/RMySQL/RMySQL.pdf)
2. [List of commands](pantz.org/software/mysql/mysqlcommands.html)
    - Do not delete, add or join items from ensembl ONLY select.
    - In general be carful with mysql commands
3. [R-bloggers](https://www.r-bloggers.com/2011/08/mysql-and-r/) - Good?
