---
title: "Reading from HDF5"
author: "MCCurcio"
date: "12/21/2020"
output: html_document
---

# Getting & Cleaning Data - Week 2
## Reading from HDF5

https://www.coursera.org/learn/data-cleaning/lecture/q7OsM/reading-from-hdf5

### HDF5
- HDF = hierarchical data format, data in groupings
- Used for large datasets and meta-data
- Supports a wide range of data types
- http://hdfgroup.org

HDF5 is installed via Bioconductor

- With R version 3.5 or greater, install Bioconductor packages using BiocManager; see 

See:
**https://bioconductor.org/install**

```{r eval=FALSE, include=FALSE}
source("http://bioconductor.org/biocLite.R")
biocLite("rhdf5")
```

Install Bioconductor Packages

To install core packages, type the following in an R command window:
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
```

Install specific packages, e.g., “rhdf5”, with

```{r}
BiocManager::install("rhdf5")
```

The lecture in the coursera course in modeled on the rhdf5 tutorial, found at: rhjdf5/vignette/pdf

```{r}
library(rhdf5)
created = h5createFile("example.h5")
created
```


### Creating Groups insdie hdf5 file
```{r}
created = h5createGroup("example.h5", "foo")
created = h5createGroup("example.h5", "baa")
created = h5createGroup("example.h5", "foo/foobaa")

h5ls("example.h5") # Like `ls` command in unix
```
### Writing to Groups
```{r}
A = matrix(data = 1:10, nrow = 5, ncol = 2)
h5write(A, "example.h5", "foo/A") 
# Write A to file example into folder "foo/A"

B = array(seq(0.1, 2.0, by = 0.1), dim=c(5,5,2))
attr(B, "scale") <- "liter" # We can add attributes like "liter"
h5write(B, "example.h5", "foo/foobaa/B")
h5ls("example.h5")
```

### Writing a Dataset directly
```{r}
df = data.frame(1L:5L, seq(0,1, length.out = 5),
                c("ab", "cde", "fghi", "a", "s"), 
                stringsAsFactors = FALSE)

h5write(df, "example.h5", "df")
h5ls("example.h5")
```

### REading the previously saved data
```{r}
readA = h5read("example.h5", "foo/A")
readB = h5read("example.h5", "foo/foobaa/B")
read_df = h5read("example.h5", "df")

readA
```

### Writing & Reading Chunks
```{r}
h5write(c(12, 13, 14), "example.h5", "foo/A", index = list(1:3, 1))
h5read("example.h5", "foo/A")
```























