---
title: "Initial Data Analysis"
author: "MCCurcio"
date: "1/16/2021"
output: 
    html_document:
        toc: true
        toc_float: true
        number_sections: true
        code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, 
                      message = FALSE, collapse = TRUE, 
                      tidy = TRUE, tidy.opts = list(arrow = TRUE, indent = 4))
```




# Initial Data Analysis

http://datasciencespecialization.github.io/getclean/

The files obtained from Coursera Getting and Data Cleaning will be inspected to see that they are ok.
```{r}
getwd()
```

Load File Names: \*test.txt & \*train.txt
```{r}
test_names = as.list(dir(pattern="*.txt"))

length(test_names)
#test_names

for (i in seq(1, length(test_names))){
    print(paste(i, test_names[i], sep = ":"))
}
```

## Find number and name(also a number) of subjects in test & train groups

### Testing participants

- Nine subjects were found to test participants
- Test participant titles = {2,4,9,10,12,13,18,20,24}
```{r}
subject_test <- read.table("~/Dropbox/R_exercises/Coursera_Materials/Data_cleaning/course_project/subject_test.txt", quote="\"", comment.char="", stringsAsFactors=FALSE)

dim(subject_test)
#unique(subject_test)
```

### Training participants

- 21 subjects were found to training participants
- Training participants titles = c(1,3,5,6,7,8,11,14,15,16,17,19,21,22,23,25,26,27,28,29,30)
```{r}
subject_train <- read.table("~/Dropbox/R_exercises/Coursera_Materials/Data_cleaning/course_project/subject_train.txt", quote="\"", comment.char="", stringsAsFactors=FALSE)

dim(subject_test)
#unique(subject_train)
#length(c(1,3,5,6,7,8,11,14,15,16,17,19,21,22,23,25,26,27,28,29,30))
```


```{r message=FALSE, warning=FALSE}
load_libs <- c("dplyr", "data.table")
lapply(load_libs, library, character.only = TRUE) # Load multiple packages

test_files_only <- c("body_acc_x_test.txt", 
                     "body_acc_y_test.txt",
                     "body_acc_z_test.txt")
                     # "body_gyro_x_test.txt",
                     # "body_gyro_y_test.txt",
                     # "body_gyro_z_test.txt")
                     # "total_acc_x_test.txt",
                     # "total_acc_y_test.txt",
                     # "total_acc_z_test.txt")

combined_files <- bind_cols(lapply(test_files_only, fread))

class(combined_files)
```















X_test.txt
X_train.txt

y_test.txt
y_train.txt

# foo <- function(file_names) {
#   lapply(file_names, load, environment())
#   ls()
# }
# foo()


# lapply(file_names,load,.GlobalEnv)

foo <- function(file_names) {
    lapply(file_names, load, .GlobalEnv)
    ls()
}
foo()
```




```{r}
my_packages <- c("dplyr", "data.table")
lapply(my_packages, library, character.only = TRUE)    # Load multiple packages

getwd()

files <- list.files(pattern = "*test.txt")

test_dt <- lapply(files, fread)

class(test_dt)
```


```{r}
for (i in 1:12) {
    print(i)
    print(dim(test_dt[i]))
}

```



```{r}
combined_files <- bind_cols(lapply(files, fread))


class(combined_files)

write.table(combined_files,
            "all_test.txt",
            sep = ",",
            row.names = FALSE)

body_acc_x_test <- read.table("~/Dropbox/R_exercises/Coursera_Materials/Data_cleaning/course_project/body_acc_x_test.txt", quote="\"", comment.char="", stringsAsFactors=FALSE)

dim(body_acc_x_test)


body_gyro_x_test <- read.table("~/Dropbox/R_exercises/Coursera_Materials/Data_cleaning/course_project/body_gyro_x_test.txt", quote="\"", comment.char="", stringsAsFactors=FALSE)
dim(body_gyro_x_test)
```

















