---
title: "EDA - Wk4 - Case Studies"
author: "MCCurcio"
date: "1/6/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

# EDA - Wk4

## Clustering Case Study

Q. What are you looking for?
Q.  What are priorities?
Q. Are your questions possible given your current dataset?

One goal of EDA is to produce a **rough-cut** of your work project.

### Samsung Dataset Example #1

UCI ML data repo:
[Human Activity Recognition Using Smartphones Data Set](https://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones)

#### Question: Can we tease out these 6 activities from the xyz data from the acceleramators and gyros of the samsung phones?
```{r echo=TRUE}
setwd("~/Dropbox/R_exercises/Coursera/Exploratory Data Analysis/week4")

library(readr)

activity_labels <- read_csv("UCI HAR Dataset/activity_labels.txt", col_names = FALSE)

head(activity_labels)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
y_test <- read_csv("UCI HAR Dataset/test/y_test.txt",  col_names = FALSE)

table(y_test)
```

### PCA and Maximum Contributor(s) REDO

See: https://www.coursera.org/learn/exploratory-data-analysis/lecture/IH1Y2/clustering-case-study

It is common to find the **Left singular** vectors. 

*Example*

    svd_data <- svd(scale(data_frame))

Then plot the first few "Left Singular" Components

*Example 1st and 2nd Left Singular Vectors*

    plot(svd_data$u[,1], col = factors)
    plot(svd_data$u[,2], col = factors)

If there are many features in your experimental space one may find it important to find which feature contributes the most to the first component.

### Find Maximum Contributor

Looking at the "Right Vector" 

    plot(svd_data$v[,2])

    maxContributor <- which.max(svd_data$v[,2])
    distanceMatrix <- dist(data_frame, maxContributor) #??????
    hClustering <- hclust(distanceMatrix)
    myplclust(hClustering, lab.col = unclass(data_frame$Y_feature)

    names(data_frame)[maxContributor]

### K-Means Clustering

    kClust <- kmeans(data_frame, centers = n)
    table(kClust$cluster, data_frame$Y_feature)

Then try **more starting** values

    kClust <- kmeans(data_frame, centers = n, nstarts = 100)
    table(kClust$cluster, data_frame$Y_feature)    



## Air Pollution Case Study

- Intro: 1D, 2D plots

- First steps - hypothesis

- Data from 1999 and more recent data from 2011-12. 

**Question** Are the levels of PM25 lower now than 1999?

EPA Air Pollution Data - https://www.epa.gov/outdoor-air-quality-data/download-daily-data

```{r}
library(readr)
file1999 <- "EPA Air Pollution data/RI_Prov_PM25_1999_site250052004_data.xls"
pm25_1999_data <- read_csv(file1999, 
                           col_types = cols(Date = col_date(format = "%m/%d/%Y")))

file2019 <- "EPA Air Pollution data/RI_Prov_PM25_2019_site250052004_data.xls"
pm25_2019_data <- read_csv(file2019, 
                           col_types = cols(Date = col_date(format = "%m/%d/%Y")))

head(pm25_1999_data)
```


### Keep Only Columns 1,4,5,7
```{r}
pm25_1999_data <- subset(pm25_1999_data, select = c(1, 5, 7))
colnames(pm25_1999_data)[1:3] <- c("Date", "PM25", "Daily.AQI")

pm25_2019_data <- subset(pm25_2019_data, select = c(1, 5, 7))
colnames(pm25_2019_data)[1:3] <- c("Date", "PM25", "Daily.AQI")

str(pm25_1999_data)
```

### Structures
```{r}
str(pm25_2019_data)
```

### Summaries
```{r}
summary(pm25_1999_data)
```

```{r}
summary(pm25_2019_data)
```

**Question** Are there any missing values? Is this important?

### Boxplots
```{r}
boxplot(pm25_1999_data$PM25, pm25_2019_data$PM25,
        main = "PM 2.5 for 1999 vs 2019 in PVD",
        ylim = c(0, 40))
```

**Question** Is there skew to the data?

**Question** Are there any negative values or NAs?

### Histograms
```{r}
hist(pm25_1999_data$PM25)
```

```{r}
hist(pm25_2019_data$PM25)
```


### Plots of Date vs PM2.5
```{r}
plot(x = pm25_2019_data$Date,
     y = pm25_2019_data$PM25,
     main = "Yr 2019 - PM 2.5 Vs Date",
     ylim = c(0, 40))
```

```{r}
plot(x = pm25_1999_data$Date,
     y = pm25_1999_data$PM25,
     main = "Yr 1999 - PM 2.5 Vs Date",
     ylim = c(0, 40))
``` 


### Graphing Time Series, Side by Side 
```{r}
par(mfrow = c(1,2))
# 1999
plot(x = pm25_1999_data$Date,
     y = pm25_1999_data$PM25,
     main = "Yr 1999 - PM 2.5 Vs Date",
     ylim = c(0, 40))
abline(h = median(pm25_1999_data$PM25), col = 2, lwd = 2)
# 2019
plot(x = pm25_2019_data$Date,
     y = pm25_2019_data$PM25,
     main = "Yr 2019 - PM 2.5 Vs Date",
     ylim = c(0, 40))
abline(h = median(pm25_1999_data$PM25), col = 2, lwd = 2)

```


**Example Code from R. Peng**

    mean0 <- with(pm0, tapply(sample.value, state.code, mean, na.rm = T))





















