---
title: "EDA-Coursera-wk2"
author: "MCCurcio"
date: "12/31/2020"
output: html_document
---

# EDA-Coursera-wk2

## Lattice Plotting System (part 1)

[Lattice1](https://www.coursera.org/learn/exploratory-data-analysis/lecture/ICqSb/lattice-plotting-system-part-1)


### xyplot, bwplot, levelplot, grid graphics, stripplot, dotplot, splom, contourplot, histograms

Lattice requires one to make the plot "all at once."


### Simple xyplot

- `xyplot(y ~ x | f * g, data)`

Example #1
```{r}
library(lattice)
library(datasets)

xyplot(Ozone ~ Wind, data = airquality)
```

### 5 Panel factor xyplot
```{r}
airquality <- transform(airquality, 
                        Month = factor(Month))
xyplot(Ozone ~ Wind | Month, 
       data = airquality, 
       layout = c(5,1))
# Plots May to Sept by factor(month)
```

Lattice Behavior
- Lattice returns an object of the class **trellis**
- auto-prints 

### Saving plots to workspace
```{r}
p <- xyplot(Ozone ~ Wind, data = airquality) # Nothing prints
print(p) # Now it prints
```

## Lattice Plotting System (part 2)



```{r}
set.seed(100)
x <- rnorm(100)
f <- rep(0:1, each = 50)
y <- x + f - f * x + rnorm(100, sd = 0.5)
f <- factor(f, labels = c("Grupo1", "Grupo2"))
xyplot(y ~ x | f, layout = c(2,1)) # plot 2 cols, 1 row
```

### Plots w/ median lines
```{r}
xyplot(y ~ x | f, panel = function(x,y, ...) {
    panel.xyplot(x,y, ...) # 1st call the default function for xyplot
    panel.abline(h = median(y), lty = 2) # Adds horizontal line at median
})
```

### Plots w/ regression lines
```{r}
xyplot(y ~ x | f, panel = function(x,y, ...) {
    panel.xyplot(x,y, ...) # 1st call the default function for xyplot
    panel.lmline(x, y, col = 3) # Adds regression lines
})
```

Example:
See:
Mouse Allergen & Asthma Cohort Study
Ahluwalia etal, J. Allergy & Clinic. Immun., 2013

Summary:

- Lattice req. one function call
- Margins and Spacing are automatics and default values usually work
- Lattice plots can be good for conditioning plots

## ggplot2 part 1

[part 1](https://www.coursera.org/learn/exploratory-data-analysis/lecture/UlFGr/ggplot2-part-1)

gg = "grammar of graphics" by Leland Wilkinson, written by Hadley W as a graduate stud.

See: http://ggplot2.org for documentation

- the grammar of graphics tells us that we have **mapping** from data to **aesthetic** attributes (color, shape, size) of **geometry** (points, lines, bars) drawn on a coordinate system.

### Basic: qplot()

```{r}
library(ggplot2)
head(mpg)
```
```{r}
str(mpg)
```
### Simple ggplot2 example
```{r}
qplot(x  =displ, y = hwy, data = mpg)
```

### Modifying Aesthetics
```{r}
qplot(x = displ, y = hwy, data = mpg, color = drv)
```

### Adding a `geom` (Geometry)

```{r}
qplot(x = displ, y = hwy, data = mpg, geom = c("point", "smooth"))
```

### Histograms
```{r}
qplot(hwy, data = mpg, fill = drv, bins = 40)
```

### Facets
```{r}
qplot(displ, hwy, data = mpg, facets = .~drv)
```
```{r}
qplot(hwy, data = mpg, facets = drv ~., binwidth = 2)
```

Example from R.Peng
- Mouse Allergen & Asthma study
- Baltimore children (5-17)
- Study indoor enviro and its relationship w asthma morbidity
- See: http://goo.gl/Wqe9j8
- Partial data currently available publically

"simulated" [maasc dataset](https://github.com/lupok2001/datasciencecoursera/blob/master/maacs.Rda)

```{r}
setwd("~/Dropbox/R_exercises/Coursera/Exploratory Data Analysis/week2")

load("maacs.Rda")
str(maacs)
```
```{r}
qplot(log(eno), data = maacs)
```

```{r}
qplot(log(eno), data = maacs, fill = mopos)
```

### Density Smooth Plots
```{r message=FALSE, warning=FALSE}
qplot(log(eno), data = maacs, geom = "density")
```

### Density Smooth Plots w/ Color
```{r message=FALSE, warning=FALSE}
qplot(log(eno), data = maacs, geom = "density", color = mopos)
```

### Scatter plot1: eNO vs. PM25
```{r message=FALSE, warning=FALSE}
qplot(log(pm25), log(eno), data = maacs)
```

### Scatter plot2: eNO vs. PM25
```{r message=FALSE, warning=FALSE}
qplot(log(pm25), log(eno), data = maacs, shape = mopos)
```

### Scatter plot3: eNO vs. PM25
```{r}
qplot(log(pm25), log(eno), data = maacs, color = mopos)
```


### Scatter plots w Smoothing and Linear Modeling

```{r message=FALSE, warning=FALSE}
qplot(log(pm25), log(eno), data = maacs, color = mopos) +
      geom_smooth(method = "lm")
```

### Splitting data with facets
```{r message=FALSE, warning=FALSE}
qplot(log(pm25), 
      log(eno), 
      data = maacs, 
      color = mopos,
      facets = . ~ mopos) +
      geom_smooth(method = "lm")
```

**NOTE: For ggplot2; It is difficult to go against the grain or customize(so don't bother) OR learn/use full ggplo2 methods with more complicated methods.**

## ggplot2 (part 3)

Fundamentals only
- Plots are built up in layers
- data frame
- aesthetic mappings: color, size
- geoms: points, lines and shapes
- facets: for conditional plots
- stats: transformations, i.e. binning, quantiles, smoothing
- scales: factors like m/f
- coordinate system

### Basic plotting
```{r message=FALSE, warning=FALSE}
qplot(log(pm25), 
      NocturnalSympt, 
      data = maacs, 
      facets = .~bmicat, 
      geom = c("point", "smooth"), 
      method = "lm")
```

### Building up layers
```{r}
head(maacs[1:3,1:3])
```
```{r}
g <- ggplot(maacs, aes(log(pm25), NocturnalSympt))
print(g)
summary(g)
```
```{r message=FALSE, warning=FALSE}
p <- g + geom_point()
print(p)
```

**OR**

```{r message=FALSE, warning=FALSE}
g + geom_point()
```

## ggplot2 (part 4)

### Adding More Layers: Smoothing

```{r message=FALSE, warning=FALSE}
g + geom_point() +
    geom_smooth()
```
```{r message=FALSE, warning=FALSE}
g + geom_point() +
    geom_smooth(method = "lm")
```

### Adding Facets - Conditional Vars
```{r message=FALSE, warning=FALSE}
g + geom_point() +
    facet_grid(. ~ bmicat) +
    geom_smooth(method = "lm")
```

### Annotations
- xlab(), ylab, labs, ggtitle
- each geom function has its own options too
- For global issues USE: **theme()**
    - `theme(legend.position = "none")
- Two std **themes** include
    - theme_gray()
    - theme_bw()

### Modifying Aesthetics (colors)
```{r message=FALSE, warning=FALSE}
g + geom_point(color = "lightblue", size = 4, alpha = 0.75) +
    geom_smooth(method = "lm")
```

```{r message=FALSE, warning=FALSE}
g + geom_point(aes(color = bmicat), size = 4, alpha = 0.5) +
    geom_smooth(method = "lm")
```

### Mod-ing Labels AND Smoothing
```{r message=FALSE, warning=FALSE}
g + geom_point(aes(color = bmicat), size = 3, alpha = 0.5) + 
                   labs(title = "MAACS Study") + 
                   labs(x = "log PM25",
                        y = "Nocturnal Symptons") +
    geom_smooth(method = "lm", se = TRUE)
```

### Changing Themes
```{r message=FALSE, warning=FALSE}
g + geom_point(aes(color = bmicat), size = 4, alpha = 0.5) +
    theme_bw(base_family = "Times") +
    geom_smooth(method = "lm")
```

## ggplot2 (part 5)

### Notes about the Axis Limits

#### Base Plot
```{r}
testdata <- data.frame(x = 1:100, y = rnorm(100))
testdata[50,2] <- 100 # Giant Outlier
plot(testdata$x, testdata$y, type = "l")
```

#### GGPlot2
```{r}
g <- ggplot(testdata, aes(x=x, y=y)) +
     geom_line()
print(g)
```

### WRONG Way
This subsets the data to ONLY include data within the limits NOT using the 50,2 datapoint.
```{r}
g + ylim(-3, 3)
```

### Better Way
This shows point including outlier with lines proceeding outside the chart limits
```{r}
g + coord_cartesian(ylim = c(-3, 3))
```

```{r}
g <- ggplot(mpg,aes(x=displ,y=hwy,color=factor(year)))
```

```{r}
g + geom_point()+facet_grid(drv~cyl, margins = TRUE)
```

```{r}
g + geom_point()+facet_grid(drv~cyl, margins = TRUE)+geom_smooth(method="lm", se=FALSE,size=2, color="black")
```

```{r}
g + geom_point()+facet_grid(drv~cyl, margins = TRUE)+geom_smooth(method="lm", se=FALSE,size=2, color="black")+labs(x = "Displacement", y= "Highway Mileage", title = "Swirl Rules!")
```



### An Other Example by R.Peng
Cutting N.O. into **RANGES**
```{r}
# Calculate the deciles of the data
cutpoints <- quantile(maacs$logno2_new, seq(0,1, length = 4), na.rm = TRUE)

# Cut the data into deciles
maacs$no_decile <- cut(maacs$logno2_new, cutpoints)

# See the levels of the newly created factor var
levels(maacs$no_decile)
```

```{r message=FALSE, warning=FALSE}
# Setup ggplot with data frame
g <- ggplot(maacs, aes(log(pm25), NocturnalSympt))

# Adding Layers
g + geom_point(alpha = 1/3) +
    facet_wrap(bmicat ~ no_decile, nrow = 2, ncol = 4) +
    geom_smooth(method = "lm", se = FALSE, col = 2) +
    theme_bw(base_family = "Times", base_size = 12) +
    labs(x = "log PM25") +
    labs(y = "Nocturnal Symptons") +
    labs(title = "MAACS Study")
```

## Practical R Exercises in swirl

### Practical R Exercises in swirl Part 2
    - Duration: 10 minutes
1. swirl Lesson 1: Lattice Plotting System
    - Duration: ~3 hours
2. swirl Lesson 2: Working with Colors
    - Duration: ~3 hours
3. swirl Lesson 3: GGPlot2 Part1
    - Duration: ~3 hours
4. swirl Lesson 4: GGPlot2 Part2
    - Duration: ~3 hours
5. swirl Lesson 5: GGPlot2 Extras
    - Duration: ~3 hours
    
## EDA Lesson 6: Lattice Plotting System


```{r}
library(lattice)
library(ggplot2)
```

Example: `xyplot(y ~ x | f * g, data)`

```{r}
head(airquality)

xyplot(Ozone ~ Wind, data = airquality)
```

```{r}
xyplot(Ozone ~ Wind, data = airquality, col = "red", pch = 8, main = "Big Apple Data")
```

```{r}
xyplot(Ozone ~ Wind | as.factor(Month), data = airquality, layout = c(5,1))
```


```{r}
xyplot(Ozone ~ Wind | Month, data = airquality, layout = c(5,1))
```

```{r}
p <- xyplot(Ozone~Wind,data=airquality)
print(p)
```


```{r}
names(p) # 45 named properties
```


```{r}
p[["formula"]]
```


```{r}
p[["x.limits"]]
```

```{r}
table(f)
```


```{r}
xyplot(y~x|f, layout = c(2,1))
```

The panel function has 3 arguments, x, y and ... .
```{r}
p <- xyplot(y ~ x | f, panel = function(x, y, ...) {
  panel.xyplot(x, y, ...)  ## First call the default panel function for 'xyplot'
  panel.abline(h = median(y), lty = 2)  ## Add a horizontal line at the median
})
print(p)
invisible()
```

Diamonds
```{r}
str(diamonds)
```

```{r}
table(diamonds$color)
```

```{r}
table(diamonds$color, diamonds$cut)
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
xyplot(price~carat|color*cut,
       data=diamonds,
       strip=FALSE,
       pch=20,
       xlab=myxlab,
       ylab=myylab,
       main=mymain)
```

### Review

| True or False? Lattice plots are constructed by a series of calls to core functions.

1: True
2: False

Selection: 2

| Excellent work!

| True or False? Lattice plots are constructed with a single function call to a core lattice function (e.g.
| xyplot)

1: True
2: False

Selection: 1

| All that hard work is paying off!


| True or False? Aspects like margins and spacing are automatically handled and defaults are usually
| sufficient.

1: False
2: True

Selection: 2

| Perseverance, that's the answer.

| True or False? The lattice system is ideal for creating conditioning plots where you examine the same kind
| of plot under many different conditions.

1: True
2: False

Selection: 1

| You are amazing!

| True or False? The lattice system, like the base plotting system, returns a trellis plot object.

1: True
2: False

Selection: 2

| That's a job well done!

| True or False?  Panel functions can NEVER be customized to modify what is plotted in each of the plot
| panels.

1: False
2: True

Selection: 1

| You nailed it! Good job!

| True or False?  Lattice plots can display at most 20 panels in a single plot.

1: False
2: True

Selection: 1


## EDA Lesson 2: Working with Colors

[04_ExploratoryAnalysis/Colors](https://github.com/DataScienceSpecialization/courses/)

```{r}
library(jpeg)
library(RColorBrewer)
library(datasets)

colors()

sample(colors(),10)

pal <- colorRamp(c("red","blue"))
pal(0)
pal(1)

pal(seq(0,1,len=6))
```

```{r}
p1 <- colorRampPalette(c("red","blue"))
p1(2)
p1(6)

0xcc
#hex 33 to decimal, as in 0x33=3*16+3=51
```


```{r}
p2 <- colorRampPalette(c("red","yellow"))
p2(2)
p2(10)
```

```{r eval=FALSE, include=FALSE}
showMe(p1(20))
```


```{r}
p1
```
function (n) 
{
    x <- ramp(seq.int(0, 1, length.out = n))
    if (ncol(x) == 4L) 
        rgb(x[, 1L], x[, 2L], x[, 3L], x[, 4L], maxColorValue = 255)
    else rgb(x[, 1L], x[, 2L], x[, 3L], maxColorValue = 255)
}
<bytecode: 0x559ae13e32d8>
<environment: 0x559ae17a2fe0>



rgb {grDevices} -	R Documentation - RGB Color Specification
```{r}
?rgb
```

rgb((0:15)/15, green = 0, blue = 0, names = paste("red", 0:15, sep = "."))

rgb(0, 0:12, 0, max = 255) # integer input

ramp <- colorRamp(c("red", "white"))
rgb( ramp(seq(0, 1, length = 5)), max = 255)

```{r}
p3 <- colorRampPalette(c("blue", "green"), alpha = 0.5)
p3(5)
```

```{r}
# Package ‘RColorBrewer’ loaded correctly!
plot(x, y, pch = 19, col = rgb(0,0.5,0.5))
```

Alpha = 0.3 (30%)
```{r}
plot(x, y, pch = 19, col = rgb(0,0.5,0.5, 0.3))

```

Our last topic for this lesson is the RColorBrewer Package, available on CRAN, that contains interesting and useful color palettes, of which there are **3 types, sequential, divergent, and qualitative**. Which one you would choose to use depends on your data.


```{r}
showMe <- function(cv){
  myarg <- deparse(substitute(cv))
  z<- outer( 1:20,1:20, "+")
  obj<- list( x=1:20,y=1:20,z=z )
  image(obj, col=cv, main=myarg  )
}
```

```{r}
cols <- brewer.pal(3, "BuGn")

pal <- colorRampPalette(cols)

showMe(pal(20))
```
Where:
showMe <- function(cv){
  myarg <- deparse(substitute(cv))
  z<- outer( 1:20,1:20, "+")
  obj<- list( x=1:20,y=1:20,z=z )
  image(obj, col=cv, main=myarg  )
}

```{r}
image(volcano, col = pal(20))
```


Which of the following is an R package that provides color palettes for sequential,
| categorical, and diverging data?

1: RColorVintner
2: RColorBrewer
3: RColorBluer
4: RColorStewer

Selection: 2

| Perseverance, that's the answer.

| True or False? The colorRamp and colorRampPalette functions can be used in conjunction with
| color palettes to connect data to colors.

1: False
2: True

Selection: 2

| That's correct!

| True or False?  Transparency can NEVER be used to clarify plots with many points

1: True
2: False

Selection: 2

| You are doing so well!

| True or False?  The call p7 <- colorRamp("red","blue") would work (i.e., not generate an
| error).

1: False
2: True

Selection: 2

| One more time. You can do it!

| Recall our reminders to concatenate the colors to form a single argument.

1: False
2: True

Selection: 1

| That's a job well done!

| True or False?  The function colors returns only 10 colors.

1: False
2: True

Selection: 1

| You're the best!

| Transparency is determined by which parameter of the rgb function?

1: beta
2: delta
3: alpha
4: it's all Greek to me
5: gamma

Selection: 3

| That's correct!



## Lesson 3: GGPlot2 Part1

GGPlot2_Part1. (Slides for this and other Data Science courses may be found at [github](https://github.com/DataScienceSpecialization/courses/)

See: http://ggplot2.org

```{r}
str(mpg)

qplot(displ, hwy, data=mpg)
```


```{r}
qplot(displ, hwy, data=mpg) + aes(color = drv)

# OR

qplot(y=hwy, data = mpg, color = drv)
```

```{r}
qplot(displ, hwy, data = mpg, color = drv, geom = c("point", "smooth"))
```

### Box plots with GGplot2
```{r}
qplot(drv, hwy, data=mpg, geom="boxplot")
```

Cool
```{r}
qplot(drv,hwy,data=mpg,geom="boxplot",color=manufacturer)
```

```{r}
qplot(hwy, data = mpg, fill = drv)
#`stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
```

```{r}
qplot(displ, hwy, data=mpg,facets = .~drv)
```

```{r}
qplot(hwy, data=mpg,facets = drv~., binwidth = 2)
```

 The facets argument, drv ~ ., resulted in what arrangement of facets?

1: huh?
2: 2 by 2
3: 3 by 1
4: 1 by 3

Selection: 3

| Keep working like that and you'll get there!

| Pretty good, right? Not too difficult either. Let's review what we learned!

...

| Which of the following is a basic workhorse function of ggplot2?

1: xyplot
2: hist
3: scatterplot
4: gplot
5: qplot

Selection: 5

| Your dedication is inspiring!

| Which types of plot does qplot plot?

1: histograms
2: scatterplots
3: all of the others
4: box and whisker plots

Selection: 3

| Excellent work!

| What does the gg in ggplot2 stand for?

1: grammar of graphics
2: good grief
3: goto graphics
4: good graphics

Selection: 1

| Keep working like that and you'll get there!

| True or False? The geom argument takes a string for a value.

1: False
2: True

Selection: 2

| You're the best!

| True or False? The data argument takes a string for a value.

1: False
2: True

Selection: 1

| Keep up the great work!

| True or False? The binwidth argument takes a string for a value.

1: False
2: True

Selection: 1

| Nice work!

| True or False? The user must specify x- and y-axis labels when using qplot.

1: True
2: False

Selection: 2


### GGPlot2 Part2


```{r}
qplot(displ, hwy, data = mpg, geom=c("point", "smooth"),facets=.~drv)
```

```{r}
g <- ggplot(mpg, aes(displ,hwy) ) 
g

summary(g)

g+geom_point()

(g + geom_point())+geom_smooth()
```

```{r}
g+geom_point()+geom_smooth(method="lm")
```


```{r}
g+geom_point()+geom_smooth(method="lm")+facet_grid(.~drv)
```

```{r}
g+geom_point(col="pink",size=4,alpha=1/2)
```

```{r}
g+geom_point(size=4,alpha=1/2, aes(color=drv))
```


```{r}
g+geom_point(aes(color=drv))+labs(title="Swirls Rules!")+labs(x="Displacement", y="Hwy Milage")
```

```{r}
g + geom_point(aes(color = drv),size=2,alpha=1/2) +
    geom_smooth(size=4,linetype=3,method="lm",se=FALSE)
```


```{r}
g + geom_point(aes(color = drv)) + theme_bw(base_family = "serif" )
```


## GGPlot2 Extras 

GGPlot2_Extras. (Slides for this and other Data Science courses may be found at github https://github.com/DataScienceSpecialization/courses/. If you care to use them, they must be downloaded as a zip file and viewed locally. This lesson corresponds to 04_ExploratoryAnalysis/ggplot2.)

```{r}
head(diamonds)
qplot(price, data=diamonds)
```

```{r}
qplot(price, data=diamonds, binwidth=18497/30, fill=cut)
```


```{r}
qplot(price, data=diamonds, geom = "density", color = cut)
```

```{r}
qplot(carat, price, data=diamonds)
```


```{r}
qplot(carat, price, data=diamonds, shape=cut)
```


```{r}
qplot(carat, price, data=diamonds, color=cut)
```


```{r}
qplot(carat, price, data=diamonds, color=cut)+ geom_smooth(method = "lm")
```


```{r}
qplot(carat,price,data=diamonds, color=cut, facets=.~cut) + geom_smooth(method="lm")
```

---

Which types of plot does qplot plot?

1: scatterplots
2: all of the others
3: histograms
4: box and whisker plots

Selection: 2

Excellent job!

Any and all of the above choices work; qplot is just that good. What does the gg in ggplot2 stand for?

1: good graphics
2: grammar of graphics
3: goto graphics
4: good grief

Selection: 2

Keep up the great work!

True or False? The geom argument takes a string for a value.

1: True
2: False

Selection: 1

You are amazing!

True or False? The method argument takes a string for a value.

1: True
2: False

Selection: 1

Keep working like that and you'll get there!

True or False? The binwidth argument takes a string for a value.

1: True
2: False

Selection: 2

Nice work!

True or False? The user must specify x- and y-axis labels when using qplot.

1: True
2: False

Selection: 2

That's the answer I was looking for.


### Making Quantiles

```{r}
cutpoints <- quantile(diamonds$carat, seq(0,1, length = 4), na.rm = TRUE)

cutpoints

diamonds$car2 <- cut(diamonds$carat, cutpoints)
```

```{r}
g <- ggplot(diamonds, aes(depth, price))
g + geom_point(alpha=1/3) + facet_grid(cut~car2)
```

```{r}
diamonds[myd,]

g + geom_point(alpha=1/3) + facet_grid(cut~car2) + geom_smooth(method="lm", size=1, color="red")
```


```{r}
ggplot(diamonds, aes(carat, price))+geom_boxplot()+facet_grid(.~cut)
```


