---
title: "Data Cleaning - Week 4"
author: "MCCurcio"
date: "1/14/2021"
output: html_document
---

# Getting and Cleaning Data - Week 4

## Editing Text Variables

We will use Baltimore Camera Data:
https://data.baltimorecity.gov/Transportation/Baltimore-Fixed-Speed-Cameras/dz54-2aru

```{r}
fixed_speed_cameras <- read.csv("~/Dropbox/R_exercises/Coursera_Materials/Data_cleaning/week4/Fixed_Speed_Cameras.csv", stringsAsFactors=FALSE)
#View(Fixed_Speed_Cameras)

names(fixed_speed_cameras)
```

### tolower

```{r}
tolower(names(fixed_speed_cameras))
```

### strsplit()
Good for data separated by periods ,etc.
```{r}
splitNames <- strsplit(names(fixed_speed_cameras), "\\.")
splitNames
```

### Quick Aside on "Lists"
```{r}
myList <- list(letters = c("A","B","C"), numbers=1:3, matrix=matrix(1:25, ncol =5))
myList
```

```{r}
myList[1]
```

OR
```{r}
myList$numbers
```

OR
```{r}
myList[[3]]
```

### Rectifying strsplit()
Since strsplit() returns a list then I can use list calls.

If I had a name like "location.23" then I could strsplit the word and the number and take only the word (for example).
```{r}
columnName = "location.23"

splitNames <- strsplit(columnName, "\\.")
splitNames
```

Making a function to take the columnName(s) and use only the first word would be:
```{r}
takeFirstWord <- function(x){x[1]}

sapply(splitNames, takeFirstWord)
```

```{r}
head(fixed_speed_cameras, 3)
```


### Sub
substituting characters for different words in my data
```{r}
tolower(names(fixed_speed_cameras))
```

ALSO: Sub is good at substituting "_" to to ""...
```{r}
sub("X", "xDegrees", names(fixed_speed_cameras))
```

### gsub
`gsub` is used for sub-ing multiple instances of a character string...
```{r}
testString <- "we want many spaces then turn them into underscores"
gsub(" ", "_", testString)
```

### grep() & grepl()
searching for specific names and strings
```{r}
grep("AVE", fixed_speed_cameras$address)
# Grep returns a vector of numbers for the locations of the sought phrase.
```

```{r}
table(grepl("AVE", fixed_speed_cameras$address))
# Grepl - grep logical, returns a logical vector of (T,F) for each line that has or does not have the sought phrase.
```

### Subsetting with `grep`
For example, I DO NOT want west bound cameras.
```{r}
not_west_bound <- fixed_speed_cameras[!grepl("W/B", fixed_speed_cameras$direction), ]
head(not_west_bound)
```

### Grep with "value=TRUE"

```{r}
grep("Ave", fixed_speed_cameras$street, value = TRUE)
```

### Using grep to find the Absence of field/string
```{r}
grep("Matt", fixed_speed_cameras$address)
```

### grep for counting
```{r}
length(grep("Ave", fixed_speed_cameras$street, value = TRUE))
```

### More string function in library(stringr)
```{r}
library(stringr)
nchar(fixed_speed_cameras[1,4])
```

### Substr
```{r}
substr(fixed_speed_cameras[1,4], 1,5) 
# 1st var: where to look,
# 2nd var: Start pt
# 3rd var: stop pt
```

### Paste
```{r}
paste("Matthew", "Curcio", sep = " ")
```

### Paste0
pastes strings together with NO characters separating them
```{r}
paste0("Matthew", "Curcio")
```

### Str_trim
- `str_trim()` removes whitespace from start and end of string
- str_trim(string, side = c("both", "left", "right"))
```{r}
str_trim("Mat       ")
```

### Str_squish
- `str_squish()` also reduces repeated whitespace inside a string
```{r}
str_squish("Matt      Curcio")
```

```{r}
str_squish("  String with trailing,  middle, and leading white space\t")
```

```{r}
str_squish("\n\nString with excess,  trailing and leading white   space\n\n")
#[1] "String with excess, trailing and leading white space"
```

### Rules

- Whenever possible convert to all lower case.
- Strings should be descriptive
    - Use "Diagnosis" rather than "Dx"
- Reduce duplication
- Not to have underscores or dots or white spaces in strings.
- Usaually are made into **Factors**


## Regular Expressions I

- Regex have **literal** and **meta** characters for use.
- Regex comes the need to search or express:
    - White space boundaries
    - Sets of literals
    - Beginning or end of line
    - Alternatives, TRUE or FALSE, is or is not


### Beginning of lines
- `^` for example, `^"My name"`, finds this phrase ONLY at beginning of lines.

### End of Line
- `$`, for example, `$"last words"` finds at end of line ONLY

### Sets or Lists of character to accept
- `[ ]`, for example, `[Bb] [Uu] [Ss] [Hh]` will find ANY combo of **bush** with any combo of caps or not.

### Combinations of searches

- `^[Ii] am` will only find "I am" or "i am" at the beginning of the line.
- `^[0-9]` : finds only numbers at the beginning of the line.
- `[a-zA-Z]$` : looks for any letter (caps or lower) at END of line.

- BUT caret(`^`) can also mean NOT.
    - `[^?.]$` will find "..." or "!" or blank space at end of line...

## Regular Expressions II

More metacharacters

- `"."` indicates 'all', `9.11` matches 9-11, or 9/11 or 9:11...
- `|` indicates OR `^Flood|Fire` matches those two words at the beginning of a line. Two "alternatives".
    - `good|bad|ugly`
    - `(!|?|.|,)$`
    - George( [Ww]\.)? Bush matches `George W. Bush` or `George Bush`
- `*` star indicates "repeat ANY number of times"
- `+` plus indicates "repeats One or more times"
    - `(.*)` matches (hello) or (), ...
    - `[0-9]+ (.*)names` matches any number before the word names.
- `{}` match a range like `^{1,5}[0-9]` matches one to five numbers together at beginning.
- `m,n` matches at least m but not more than n mathces
- `m` matches exactly `^m999` must have **999** at beginning not 99 or 9999.
- `m,` matches 'at least m number of matches.
- ` +([aA-Z]+) +\1+ 1` matches "blah blah" or "and and", ...
- the `*` is greedy such that it matches the longest possible phrase.
    - `^s(.*)s` matches "sitting with my sister at starbucks"
    - For LESS greedy lengths to match add `?`
    - `^[sS](.*?)s$` looks for small length such as "sitting with my sister at starbucks" returns **"sitting with my s"**

## Working with Dates

```{r}
d1 = date()
d1
```
```{r}
class(d1)
```

### Using `Sys.Date()`

```{r}
d2 = Sys.Date()
d2
```

```{r}
class(d2)
```

### formatting dates
```{r}
format(d2, "%a %b %d")
```

SEE: **man date**

### taking and formatting into date
```{r}
x = c("1Jan1960", "14Mar1975")
z = as.Date(x, "%d%b%Y")
z
```

### Time differences
```{r}
z[2] - z[1]
```

```{r}
as.numeric(z[2] - z[1])
```

### Converting to Julian
```{r}
weekdays(z[1])
```

```{r}
months(z[1])
```

### From "1960-01-01"
```{r}
julian(z[1])
```

### Lubridate
```{r}
library(lubridate)

?Sys.Date
?Sys.timezone

# Convert a number ot a date
ymd("19640829") # or any combo of 'ymd' for reformatting

ymd_hms("2015081410:20:25")

# Setting time zones
ymd_hms("20150814 10:20:25", tz="EST")
```

NOTE: Base R is slightly different
```{r}
x = dmy(c("1Jan1960", "14Mar1975"))
wday(x[1])
```

See:
- [Lubridate](https://lubridate.tidyverse.org/)
- It is best to use "POSIXct" or "POSIXlt"
https://www.rdocumentation.org/packages/tis/versions/1.38/topics/POSIXct


## Data Resources

Free data is everywhere!

- United Nations - data.un.org
- US - data.gov
- UK - data.gov.uk
- France - www.data.gouv.fr
- MORE - www.data.gov/opendatasites

- Gapminder is cool! - gapminder.org
- Survey data - asdfree.com
- Infochimps Marketplace
- Kaggle
- UCI
- Kdnuggets
- Arvix data
- Public data on Amazon Web Services
- CMU Stalib
- Stanford Large Network data

See also: Hilary Mason, Peter Skomoroch, Jeff Hammerbacher, Gregory Piatetshy-shapiro,... 

#### APIs include
 
- twitter, PLoS, rOpenSci, Google maps, figshare











